Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-12 shanghai.tex
Made a few more changes to the section on base change over finite fields.

When using geometrized characters to study characters, it is useful to understand
how geometrized characters behave under change of fields.
Suppose that $k'$ is a finite extension of $k$ of degree $n$, 
$G$ is a commutative smooth group scheme locally of finite type over $k$.
If $\cs{L} = (\gcs{L}, \mu, \phi)$ is a geometrized character of $G$ we may define
a geometrized character $\cs{L}' = (\gcs{L}, \mu, \phi_{k'})$ on 
$G_{k'}\ceq G\times_{\Spec{k}} \Spec{k'}$ by setting
\[
\phi_{k'} = \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (CS.2) for $\phi_{k'}$ 
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as taking the action $\varphi$ of $\Weil{k}$ on $\gcs{L}$ 
defined in Section~\ref{sec:category} and restricting it to the subgroup $\Weil{k'}$.

\begin{proposition}\label{prop:basechange}
 The rule $(\gcs{L}, \mu, \varphi) \mapsto (\gcs{L}, \mu, \varphi\vert_{\Weil{\Fq}})$ 
 defines a monoidal functor $\GC(G) \to \GC(G_{k'})$. 
 Moreover, the following diagram commutes, 
 where the top map is induced by $\GC(G) \to \GC(G_{k'})$ 
 and the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}[column sep=60]
\GCiso{G} \rar{\text{ext'n of scalars}} \dar{\trFrob{G}} & \GCiso{G_{k'}} \dar{\trFrob{G_{k'}}} \\
G(k)^*  \rar{\Nm} & G(k')^*
%\Hom(G(k), \EEx) \rar{\Nm} & \Hom(G(k'), \EEx).
\end{tikzcd}
\]
\end{proposition}

\begin{proof}
\Clifton{This proof is missing. Can David insert it?}
\end{proof}

@@

\section{Restriction of scalars}

In the opposite direction, there is a canonical closed immersion of $k$-schemes
$G \hookrightarrow \Res_{k'/k}(G_{k'})$ \cite[\S 7.6]{BLR}.  Pullback yields a functor
\[
\GC(\Res_{k'/k}(G_{k'})) \xrightarrow{\iota} \GC(G).
\]
\begin{proposition} \label{prop:csbe}
The following diagram commutes, where the bottom map is induced by the inclusion $G(k) \hookrightarrow G(k')$:
\[
\begin{tikzcd}
\GCiso{\Res_{k'/k}(G_{k'})} \arrow[two heads]{r}{\iota} \dar{t_{\Res_{k'/k}(G_{k'})}} & \GCiso{G} \dar{t_G} \\
\Hom(G(k'), \EEx) \arrow[two heads]{r} & \Hom(G(k), \EEx).
\end{tikzcd}
\]
\end{proposition}

Now suppose that $G'$ is an abelian group scheme over $k'$, satisfying the conditions of Theorem \ref{thm:snake}.
There is a canonical isomorphism between $G'(k')$ and $(\Res_{k'/k}G')(k)$;
we now describe the corresponding functor from
$\GC(G')$ to $\GC(\Res_{k'/k}G')$.  The base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
decomposes into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
The component corresponding to the identity element yields a natural inclusion 
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'}
\]
of $k'$-schemes.  The definition of base change yields a map of $k$-schemes
\[
(\Res_{k'/k}G')_{k'} \to \Res_{k'/k}G'.
\]
Consecutive pullback along these maps yields a functor
\[
\GC(G') \to \GC(\Res_{k'/k}G').
\]

\begin{proposition}
  Then the following diagram commutes:
\[
\begin{tikzcd}
\GCiso{G'} \dar{t_{G'}} \rar & \GCiso{\Res_{k'/k} G'} \dar{t_{\Res_{k'/k} G'}} \\
 \Hom(G'(k'), \EEx) \rar & \Hom((\Res_{k'/k}G')(k), \EEx)
\end{tikzcd}
\]
\end{proposition}

@@

\section{Base change over the local field}

We have already seen that the trace of Frobenius defines a canonical isomorphism 
$t_T : \GCiso{\GN{T}} \to \Hom_\text{ad}(T(K),\EEx)$. 
In this section we expand on the canonicity of this group isomorphism 
by showing that it is a natural transformation. We also show

@@

>>>>

When using geometrized characters to study characters, it is useful to understand
how geometrized characters behave under change of fields.
Let $k'$ be a finite extension of $k$. Then $k\to k'$ induces a group homomorphism $i_{k'/k} : G(k) \hookrightarrow G(k')$ which, in turn, gives
\[
i_{k'/k}^* : G(k')^* \to G(k)^* \qquad \text{defined by}\qquad \chi \mapsto \chi\circ i_{k'/k}.
\]
We can interpret this operation on characters in terms of geometrized characters, as follows.
%

\begin{proposition} \label{prop:csbe}
Set $G_{k'} = G\times_\Spec{k} \Spec{k'}$ and let
\[
\GC(\Res_{k'/k}(G_{k'})) \xrightarrow{\iota^*} \GC(G)
\]
be the functor obtained by pull-back along the canonical closed immersion of $k$-schemes  $\iota : G \hookrightarrow \Res_{k'/k}(G_{k'})$ \cite[\S 7.6]{BLR}. 
The following diagram commutes:
\[
\begin{tikzcd}
\GCiso{\Res_{k'/k}(G_{k'})} \arrow[two heads]{r}{\iota} \dar[swap]{t_{\Res_{k'/k}(G_{k'})}} & \GCiso{G} \dar{t_G} \\
G(k')^* \arrow[two heads]{r}{i_{k'/k}^*} & G(k)^*.
\end{tikzcd}
\]
\end{proposition}

In the opposite direction, let $\Nm_{k'/k} : G(k') \to G(k)$ be the norm map and consider the group homomorhism:
\Clifton{How is $\Nm$ defined?}
\[
\Nm_{k'/k}^* : G(k)^* \to G(k')^* \qquad \text{defined by}\qquad \chi \mapsto \chi\circ \Nm_{k'/k}.
\]
We can also interpret this operation on characters in terms of geometrized characters, as follows.

If $\cs{L} = (\gcs{L}, \mu, \phi)$ is a geometrized character of $G$ we may define
a geometrized character $\cs{L}' = (\gcs{L}, \mu, \phi_{k'})$ on 
$G_{k'}\ceq G\times_{\Spec{k}} \Spec{k'}$ by setting
\[
\phi_{k'} = \phi \circ \Frob{G}^*(\phi) \circ \cdots \circ (\Frob{G}^{n-1})^*(\phi).
\]
The commutativity of the diagram (CS.2) for $\phi_{k'}$ 
follows from the fact that $\Frob{G_{k'}} = \Frob{G}^n$.
Note that we may also think about the construction of $\phi_{k'}$ from $\phi$
as taking the action $\varphi$ of $\Weil{k}$ on $\gcs{L}$ 
defined in Section~\ref{sec:category} and restricting it to the subgroup $\Weil{k'}$.

\begin{proposition}\label{prop:basechange}
 The rule $\nu_{k'/k}: (\gcs{L}, \mu, \varphi) \mapsto (\gcs{L}, \mu, \varphi\vert_{\Weil{\Fq}})$ 
 defines a monoidal functor $\GC(G) \to \GC(G_{k'})$. 
 Moreover, the following diagram commutes, 
 where the top map is induced by $\GC(G) \to \GC(G_{k'})$ 
 and the bottom map is induced by the norm $G(k') \rightarrow G(k)$:
\[
\begin{tikzcd}[column sep=60]
\GCiso{G} \rar{\nu_{k'/k}} \dar{\trFrob{G}} & \GCiso{G_{k'}} \dar{\trFrob{G_{k'}}} \\
G(k)^*  \rar{\Nm_{k'/k}^*} & G(k')^*.
\end{tikzcd}
\]
\end{proposition}

Finally, we explain how to geometrize the canonical isomorphism between characters of $G'(k')$ and of $(\Res_{k'/k}G')(k)$.
The base change $(\Res_{k'/k}G')_{k'}$ of $\Res_{k'/k}G'$ to $k'$
decomposes into a product of copies of $G'$, indexed by elements of $\Gal(k'/k)$:
\[
(\Res_{k'/k}G')_{k'} \cong \prod_{\Gal(k'/k)} G'.
\]
The component corresponding to the identity element yields a natural inclusion 
\[
G' \hookrightarrow (\Res_{k'/k}G')_{k'}
\]
of $k'$-schemes.  The definition of base change yields a map of $k$-schemes
\[
(\Res_{k'/k}G')_{k'} \to \Res_{k'/k}G'.
\]
Consecutive pullback along these maps yields a functor
\[
\GC(G') \to \GC(\Res_{k'/k}G').
\]

\begin{proposition}
The following diagram commutes:
\[
\begin{tikzcd}
\GCiso{G'} \dar{t_{G'}} \rar & \GCiso{\Res_{k'/k} G'} \dar{t_{\Res_{k'/k} G'}} \\
 G'(k')^* \rar & (\Res_{k'/k}G')(k)^*
\end{tikzcd}
\]
\end{proposition}

@@

@@

\section{Canonicity}

We have already seen that the trace of Frobenius defines a canonical isomorphism 
$t_T : \GCiso{\GN{T}} \to \Hom_\text{ad}(T(K),\EEx)$. 
In this section we expand on the canonicity of this group isomorphism.

@@

\section{Geometrization of restriction of scalars of local fields}
